---
title: "Data Presentation"
author: "Paul Thompson"
date: "25 October 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Population Health Impacts from storms

Storms have considerable impact on public health, this analysis should outline the types of events that are harmful to population health. Results about ecenomic impact will also be revealed.

### X and Y as health risks and ecenomic factors in storm impact

# Data Processing

The storm data provided is in a good, format for the most. Whitespace is stripped for ease of referencing

```{r, cache=TRUE}
library(ggplot2)
data <- read.csv("StormData.csv.bz2", strip.white = T) 
```

Valida values for ecenomic EXP categories are K, M, and B for thousands, millions and billions. Lets start by only taking the rows that have this data 

```{r}
economicEvents <- data[data$PROPDMGEXP %in% c("K", "M", "B"),]
```

Now, for those that are K, we times by 1000, M by 1,000,000, and B by 1,000,000,000
```{r}
money <- list()
for(i in 1:nrow(economicEvents)) {
  if(economicEvents[i,]$PROPDMGEXP == "K") {
    money[i] <- economicEvents[i,]$PROPDMG * 1000
  } else if(economicEvents[i,]$PROPDMGEXP == "M") {
    money[i] <- economicEvents[i,]$PROPDMG * 1000000
  } else {
    money[i] <- economicEvents[i,]$PROPDMG * 1000000000
  }
}
```

So as we're looking into the events that are most harmful with 

```{r}
fatalities <- aggregate(data$FATALITIES, by=list(data$EVTYPE), sum)
injuries <- aggregate(data$INJURIES, by=list(data$EVTYPE), sum)

highestFatality <- fatalities[which.max(fatalities$x),][1]
highestInjury <- injuries[which.max(fatalities$x),][1]
```

We also need to find out which events have the most economic impact. 

```{r}
includedDamage <- c("CROPDMGEXP", "PROPDMGEXP", )
```

# Results

### Across the United States, which types of events are most harmful with respect to population health?

This question requires taking a look at both the events that are hamful via death and injuries. Lets summarise first by finding the events that are 

* The most deadly
* The most injury creating

Lets take a look at the top 10 in each category

```{r}
attach(fatalities)
topX <- head(fatalities[order(-x),], n=10)

ggplot(topX, aes(x=Group.1, y=x)) + 
  geom_bar(stat="identity") +
  labs(x="Event", y="Fatalities") + 
  theme(axis.text.x = element_text(angle = 45))

attach(injuries)
topXIn <- head(injuries[order(-x),], n=10)

ggplot(topXIn, aes(x=Group.1, y=x)) + 
  geom_bar(stat="identity") +
  labs(x="Event", y="Fatalities") + 
  theme(axis.text.x = element_text(angle = 45))
```

So, obviously the answer is the same for each one:

```{r}
all.equal(highestInjury, highestFatality)

highestFatality
```

So, from our data. Tornados are the worst event for population health. By a significant margin too.

